### 2023-02-24 products/manuscript/Manuscript.qmd

---
title: "MADA Final Project"
subtitle: "Trends in New HIV Diagnoses in the US, 2008-2022"
author: Nicole Luisi
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

Note: There seem to be  lot of issues getting tables to render well in Word doc format, need to come back to this. I tried 5 or so different methods and they all look terrible when in a doc vs html. Also having issues with resolving crossrefs.

```{r, echo=FALSE, message=FALSE}
# load R packages
library(here) 
library(ggplot2)
library(skimr) 
library(plotly)
library(gt)
library(knitr)
```

# Summary/Abstract

The purpose of this analysis is to analyze US trends in new HIV diagnoses over the past 10 years. Previous research by [Chapin-Bardales et al. (2017)](https://doi.org/10.1016/j.annepidem.2017.04.002) evaluated trends in AIDS diagnosis rates by race/ethnicity between 1984-2013. This analysis will focus on new HIV diagnosesbetween  2008-2022; subgroups will include region, age group, gender, and race/ethnicity. [@Chapin2017]

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source

**Data Source**:

*Note: New diagnoses data are currently available through 2020. If 2021-2022 data are not available by the time the analysis needs to be completed, the analysis will be temporarily restricted to 2012-2020.*

+ HIV New Diagnoses: [AIDSVu](https://aidsvu.org/) [@Sullivan2020]

**Analysis**:

+ Rate (per 100,000) of new HIV diagnoses by age, gender, and race/ethnicity (national, regional)
+ Rate ratios for new HIV diagnoses by age, gender, and race/ethnicity (national, regional)
+ Estimated annual percent change (EAPC) for incidence
  - Calculated using [Joinpoint Regression Software (R-Callable Command-Line Version 4.9.0.0)](https://surveillance.cancer.gov/help/joinpoint/tech-help/citation)
+ Other modeling TBD based on course modules

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Data aquisition
Data files were downloaded from [AIDSVu](https://aidsvu.org/) and placed in the `data/raw_data/AIDSVu` subfolder. The new diagnosis files consist of one xlsx file per year for each level (national, regional, etc.).

## Data import and cleaning
The `processingcode.R` file includes the following steps:
+ Read in xlsx data files and combine all years for each level
+ Subset variables of interest and rename variables so all levels are consistent 
+ View/explore data
+ Calculate new rate ratio variables for subgroups 
+ Prepare combined national and regional file with all variables

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

@tbl-natsummarytable shows a full summary of the National data.

```{r}
#| label: tbl-natsummarytable
#| tbl-cap: "National data summary table."
#| echo: FALSE
alldata<-readRDS(here("data", "processed_data", "allnewdx.rds"))
national<-alldata %>% filter(Geo %in% c("United States"))
national %>% gt::gt() %>% 
  gt::tab_header(title = "National Data") %>%
  gt::tab_style(style = list(gt::cell_fill(color = "#7cb9e8"),
                             gt::cell_text(weight = "bold")),
                locations = gt::cells_body(columns = Year))
```
@fig-natplots shows exploratory plots for the national rates by subgroup.

```{r}
#| label: fig-natplots
#| fig-cap: "National rates by subgroup"
#| echo: FALSE
national<-alldata %>% filter(Geo %in% c("United States"))
natfig <- plot_ly(national, x = ~Year, y = ~Overall_Rate, name = "Overall", type="scatter", mode = "lines", line = list(width = 3, dash = "dash", color = "gray")) 
natfig <- natfig %>% add_trace(y = ~Male_Rate, name = "Male", mode = 'lines', line = list(width = 3, dash = "solid", color = "blue"))
natfig <- natfig %>% add_trace(y = ~Female_Rate, name = "Female", mode = 'lines', line = list(width = 3, dash = "solid", color = "green"))
natfig %>% layout(title="National New Diagnosis Rate by Sex", legend=list(title=list(text='Sex')), 
                  xaxis = list(title = 'Year'),
                  yaxis = list(title = 'Rate per 100,000'))

natfig <- plot_ly(national, x = ~Year, y = ~Overall_Rate, name = "Overall", type="scatter", mode = "lines", line = list(width = 3, dash = "dash", color = "gray")) 
natfig <- natfig %>% add_trace(y = ~Black_Rate, name = "Black", mode = 'lines', line = list(width = 3, dash = "solid", color = "blue"))
natfig <- natfig %>% add_trace(y = ~White_Rate, name = "White", mode = 'lines', line = list(width = 3, dash = "solid", color = "darkgreen"))
natfig <- natfig %>% add_trace(y = ~Hispanic_Rate, name = "Hispanic", mode = 'lines', line = list(width = 3, dash = "solid", color = "darkred"))
natfig <- natfig %>% add_trace(y = ~Asian_Rate, name = "Asian", mode = 'lines', line = list(width = 3, dash = "solid", color = "cornflowerblue"))
natfig <- natfig %>% add_trace(y = ~AIAN_Rate, name = "AIAN", mode = 'lines', line = list(width = 3, dash = "solid", color = "brown1"))
natfig <- natfig %>% add_trace(y = ~MultRace_Rate, name = "Multiracial", mode = 'lines', line = list(width = 3, dash = "solid", color = "purple"))
natfig <- natfig %>% add_trace(y = ~NHPI_Rate, name = "NHPI", mode = 'lines', line = list(width = 3, dash = "solid", color = "magenta"))
natfig %>% layout(title="National New Diagnosis Rate by Race/Ethnicity", legend=list(title=list(text='Race/Ethnicity')), 
                  xaxis = list(title = 'Year'),
                  yaxis = list(title = 'Rate per 100,000'))

natfig <- plot_ly(national, x = ~Year, y = ~Overall_Rate, name = "Overall", type="scatter", mode = "lines", line = list(width = 3, dash = "dash", color = "gray")) 
natfig <- natfig %>% add_trace(y = ~Age13t24_Rate, name = "Age 13-24", mode = 'lines', line = list(width = 3, dash = "solid", color = "blue"))
natfig <- natfig %>% add_trace(y = ~Age25t34_Rate, name = "Age 25-34", mode = 'lines', line = list(width = 3, dash = "solid", color = "darkgreen"))
natfig <- natfig %>% add_trace(y = ~Age35t44_Rate, name = "Age 35-44e", mode = 'lines', line = list(width = 3, dash = "solid", color = "darkred"))
natfig <- natfig %>% add_trace(y = ~Age45t54_Rate, name = "Age 45-54", mode = 'lines', line = list(width = 3, dash = "solid", color = "cornflowerblue"))
natfig <- natfig %>% add_trace(y = ~Age55p_Rate, name = "Age 55+", mode = 'lines', line = list(width = 3, dash = "solid", color = "orange"))
natfig %>% layout(title="National New Diagnosis Rate by Age", legend=list(title=list(text='Age')), 
                  xaxis = list(title = 'Year'),
                  yaxis = list(title = 'Rate per 100,000'))
```

@tbl-natratios shows a summary of selected rate ratios for race/ethnicity and sex.

```{r}
#| label: tbl-natratios
#| tbl-cap: "National rate ratios by select subgroups"
#| echo: FALSE
national<-alldata %>% filter(Geo %in% c("United States"))
national_sub <- national %>% dplyr::select("Year", "Black_Rate", "White_Rate", "Hispanic_Rate", "black_white_rateratio", "hispanic_white_rateratio", "Male_Rate", "Female_Rate", "female_male_rateratio")
national_sub %>% gt::gt() %>% 
  gt::tab_header(title = "Rate Ratios") %>%
  gt::tab_style(style = list(gt::cell_fill(color = "#7cb9e8"),
                             gt::cell_text(weight = "bold")),
                locations = gt::cells_body(columns = Year))%>%
  gt::tab_style(style = list(gt::cell_fill(color = "#F1EB9C"),
                             gt::cell_text(weight = "bold")), 
                locations = gt::cells_body(columns = c(black_white_rateratio,hispanic_white_rateratio,female_male_rateratio)))
```

@tbl-regsummarytable shows a full summary of the Regional data.

```{r}
#| label: tbl-regsummarytable
#| tbl-cap: "Regional data summary table."
#| echo: FALSE
alldata<-readRDS(here("data", "processed_data", "allnewdx.rds"))
regional<-alldata %>% filter(Geo %in% c("Northeast", "Midwest", "South", "West"))
regional %>% gt::gt() %>% 
  gt::tab_header(title = "Regional Data") %>%
  gt::tab_style(style = list(gt::cell_fill(color = "#7cb9e8"),
                             gt::cell_text(weight = "bold")),
                locations = gt::cells_body(columns = Year))
```

@fig-regplots shows an exploratory plot for regional overall new diagnosis rates with female-male rate ratios.

```{r}
#| label: fig-regplots
#| tbl-cap: "Regional plots"
#| echo: FALSE
alldata<-readRDS(here("data", "processed_data", "allnewdx.rds"))
regional<-alldata %>% filter(Geo %in% c("Northeast", "Midwest", "South", "West"))
ggplot(regional, aes(x=Year, y=log(Overall_Rate), size=log(female_male_rateratio), color=Geo)) +
    geom_point(alpha=0.7) + scale_x_continuous(breaks=c(2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  ggtitle("Regional New Diagnosis Rates with Female-Male Rate Ratios") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Year") + ylab("Log of Overall NewDx Rate")
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by sex."
#| echo: FALSE
#knitr::include_graphics(here("results","height_weight_stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
#resulttable2 = readRDS(here("results","resulttable2.rds"))
#knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [] discusses types of analyses.

These papers [] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References