### 2023-04-21 code/analysis_code/analysis.qmd
Submit Part 5

---
title: "Analysis script"
author: "Nicole Luisi"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

```{r, echo=FALSE, message=FALSE}
# load R packages
library(here) 
library(readxl)
library(ggplot2)
library(skimr) 
library(plotly)
library(gt)
library(knitr)
library(dplyr)
library(kableExtra)
library(gridExtra)
library(ggpubr)
library(janitor)
library(tidymodels)
library(dotwhisker)
library(tidyverse)
library(caret)
```

# Load clean data

```{r}
alldata<-readRDS(here("data", "processed_data", "allnewdx.rds"))
```

```{r, echo=FALSE, message=FALSE}
# Plot of US rates by sex, age, race/ethnicity 
national<-alldata %>% filter(Geo %in% c("United States"))
natfig_s <- plot_ly(national, x = ~Year, y = ~Overall_Rate, name = "Overall", type="scatter", mode = "lines", line = list(width = 3, dash = "dash", color = "gray")) 
natfig_s <- natfig_s %>% add_trace(y = ~Male_Rate, name = "Male", mode = 'lines', line = list(width = 3, dash = "solid", color = "blue"))
natfig_s <- natfig_s %>% add_trace(y = ~Female_Rate, name = "Female", mode = 'lines', line = list(width = 3, dash = "solid", color = "green"))
natfig_s <- natfig_s %>% layout(title="National New Diagnosis Rate by Sex", legend=list(title=list(text='Sex')), 
                  xaxis = list(title = 'Year'),
                  yaxis = list(title = 'Rate per 100,000'))

natfig_r <- plot_ly(national, x = ~Year, y = ~Overall_Rate, name = "Overall", type="scatter", mode = "lines", line = list(width = 3, dash = "dash", color = "gray")) 
natfig_r <- natfig_r %>% add_trace(y = ~Black_Rate, name = "Black", mode = 'lines', line = list(width = 3, dash = "solid", color = "blue"))
natfig_r <- natfig_r %>% add_trace(y = ~White_Rate, name = "White", mode = 'lines', line = list(width = 3, dash = "solid", color = "darkgreen"))
natfig_r <- natfig_r %>% add_trace(y = ~Hispanic_Rate, name = "Hispanic", mode = 'lines', line = list(width = 3, dash = "solid", color = "darkred"))
natfig_r <- natfig_r %>% add_trace(y = ~Asian_Rate, name = "Asian", mode = 'lines', line = list(width = 3, dash = "solid", color = "cornflowerblue"))
natfig_r <- natfig_r %>% add_trace(y = ~AIAN_Rate, name = "AIAN", mode = 'lines', line = list(width = 3, dash = "solid", color = "brown1"))
natfig_r <- natfig_r %>% add_trace(y = ~MultRace_Rate, name = "Multiracial", mode = 'lines', line = list(width = 3, dash = "solid", color = "purple"))
natfig_r <- natfig_r %>% add_trace(y = ~NHPI_Rate, name = "NHPI", mode = 'lines', line = list(width = 3, dash = "solid", color = "magenta"))
natfig_r <- natfig_r %>% layout(title="National New Diagnosis Rate by Race/Ethnicity", legend=list(title=list(text='Race/Ethnicity')), 
                  xaxis = list(title = 'Year'),
                  yaxis = list(title = 'Rate per 100,000'))

natfig_a <- plot_ly(national, x = ~Year, y = ~Overall_Rate, name = "Overall", type="scatter", mode = "lines", line = list(width = 3, dash = "dash", color = "gray")) 
natfig_a <- natfig_a %>% add_trace(y = ~Age13t24_Rate, name = "Age 13-24", mode = 'lines', line = list(width = 3, dash = "solid", color = "blue"))
natfig_a <- natfig_a %>% add_trace(y = ~Age25t34_Rate, name = "Age 25-34", mode = 'lines', line = list(width = 3, dash = "solid", color = "darkgreen"))
natfig_a <- natfig_a %>% add_trace(y = ~Age35t44_Rate, name = "Age 35-44", mode = 'lines', line = list(width = 3, dash = "solid", color = "darkred"))
natfig_a <- natfig_a %>% add_trace(y = ~Age45t54_Rate, name = "Age 45-54", mode = 'lines', line = list(width = 3, dash = "solid", color = "cornflowerblue"))
natfig_a <- natfig_a %>% add_trace(y = ~Age55p_Rate, name = "Age 55+", mode = 'lines', line = list(width = 3, dash = "solid", color = "orange"))
natfig_a <- natfig_a %>% layout(title="National New Diagnosis Rate by Age", legend=list(title=list(text='Age')), 
                  xaxis = list(title = 'Year'),
                  yaxis = list(title = 'Rate per 100,000'))

natfig_s
plotly::export(p=natfig_s, file = here("results", "US_rates_bysex.png"))
natfig_a
plotly::export(p=natfig_a, file = here("results", "US_rates_byage.png"))
natfig_r
plotly::export(p=natfig_r, file = here("results", "US_rates_byrace.png"))
```

# Load data from external Joinpoint program

```{r, echo=FALSE, message=FALSE}
# Read in data from external Joinpoint program

# Function to pull names of sheets in excel files (just applying to one since all are the same)

multiplesheets <- function(fname) {
  sheets <- readxl::excel_sheets(fname)
  tibble <- lapply(sheets, function(x) readxl::read_excel(fname, sheet = x))
  data_frame <- lapply(tibble, as.data.frame)
  names(data_frame) <- sheets
  print(data_frame)
}
path <- here("data", "processed_data", "djp_overall.xlsx")
multiplesheets(path)

# Overall
jpoutput_overall_aapc <- read_excel(here("data", "processed_data", "djp_overall.xlsx"), sheet="AAPCs")
names(jpoutput_overall_aapc)<- c("geo", "joinpoint_model", "AAPC_Index", "start", "end", "AAPC", "AAPC_lCI", "AAPC_uCI","significant", "test_statstic", "p-value")
jpoutput_overall_apc <- read_excel(here("data", "processed_data", "djp_overall.xlsx"), sheet="APCs")
names(jpoutput_overall_apc)<-c("Geo","Model","Segment","Segment Start","Segment End","APC","APC 95% LCL","APC 95% UCL","APC Significant","Test Statistic","P-Value")

# Race
jpoutput_race_aapc <- read_excel(here("data", "processed_data", "djp_race.xlsx"), sheet="AAPCs")
names(jpoutput_race_aapc)<- c("type","geo", "joinpoint_model", "AAPC_Index", "start", "end", "AAPC", "AAPC_lCI", "AAPC_uCI","significant", "test_statstic", "p-value")
jpoutput_race_data <- read_excel(here("data", "processed_data", "djp_race.xlsx"), sheet="Data")
names(jpoutput_race_data)<- c("type","geo", "year", "rate", "jp", "flag")
jpoutput_race_apc <- read_excel(here("data", "processed_data", "djp_race.xlsx"), sheet="APCs")
names(jpoutput_race_apc)<-c("Type","Geo","Model","Segment","Segment Start","Segment End","APC","APC 95% LCL","APC 95% UCL","APC Significant","Test Statistic","P-Value")

# Race Ratios
jpoutput_raceratios_aapc <- read_excel(here("data", "processed_data", "djp_raceratios.xlsx"), sheet="AAPCs")
names(jpoutput_raceratios_aapc)<- c("type","geo", "joinpoint_model", "AAPC_Index", "start", "end", "AAPC", "AAPC_lCI", "AAPC_uCI","significant", "test_statstic", "p-value")
jpoutput_raceratio_data <- read_excel(here("data", "processed_data", "djp_raceratios.xlsx"), sheet="Data")
names(jpoutput_raceratio_data)<- c("type","geo", "year", "rate", "jp", "flag")
jpoutput_raceratio_apc <- read_excel(here("data", "processed_data", "djp_raceratios.xlsx"), sheet="APCs")
names(jpoutput_raceratio_apc)<-c("Type","Geo","Model","Segment","Segment Start","Segment End","APC","APC 95% LCL","APC 95% UCL","APC Significant","Test Statistic","P-Value")

# Age
jpoutput_age_aapc <- read_excel(here("data", "processed_data", "djp_age.xlsx"), sheet="AAPCs")
names(jpoutput_age_aapc)<- c("type","geo", "joinpoint_model", "AAPC_Index", "start", "end", "AAPC", "AAPC_lCI", "AAPC_uCI","significant", "test_statstic", "p-value")
jpoutput_age_apc <- read_excel(here("data", "processed_data", "djp_age.xlsx"), sheet="APCs")
names(jpoutput_age_apc)<-c("Type","Geo","Model","Segment","Segment Start","Segment End","APC","APC 95% LCL","APC 95% UCL","APC Significant","Test Statistic","P-Value")

# Sex
jpoutput_sex_aapc <- read_excel(here("data", "processed_data", "djp_sex.xlsx"), sheet="AAPCs")
names(jpoutput_sex_aapc)<- c("type","geo", "joinpoint_model", "AAPC_Index", "start", "end", "AAPC", "AAPC_lCI", "AAPC_uCI","significant", "test_statstic", "p-value")
jpoutput_sex_apc <- read_excel(here("data", "processed_data", "djp_sex.xlsx"), sheet="APCs")
names(jpoutput_sex_apc)<-c("Type","Geo","Model","Segment","Segment Start","Segment End","APC","APC 95% LCL","APC 95% UCL","APC Significant","Test Statistic","P-Value")
```

## Table of AAPCs

```{r, echo=FALSE, message=FALSE}
jpoutput_overall_aapc$type<-"Overall"
jpoutput_overall_aapc <- jpoutput_overall_aapc[, c(12, 1:11)]
all_aapc<-rbind(jpoutput_overall_aapc,jpoutput_sex_aapc,jpoutput_age_aapc,jpoutput_race_aapc,jpoutput_raceratios_aapc) %>% select("type","geo", "AAPC", "AAPC_lCI", "AAPC_uCI", "p-value")

kable(all_aapc,caption = "All AAPC Results", 
      col.names=c("Stratification", "Geographic Area", "AAPC", "AAPC_lCI", "AAPC_uCI", "p-value"),      align="rc") %>% 
  row_spec(row = 0, color = "white", background = "#004987") %>% kable_styling(full_width=T, bootstrap_options = "striped", font_size=10) %>% 
  pack_rows("Overall", 1, 5) %>%
  pack_rows("Female", 6, 10) %>%
  pack_rows("Male", 11, 15) %>%
  pack_rows("Age 13-24", 16, 20) %>%
  pack_rows("Age 25-34", 21, 25) %>%
  pack_rows("Age 35-44", 26, 30)  %>%
  pack_rows("Age 45-54", 31, 35)  %>%
  pack_rows("Age 55+", 36, 40)  %>%
  pack_rows("American Indian/Alaska Native", 41, 44) %>%
  pack_rows("Asian", 45, 49) %>%
  pack_rows("Black", 50, 54) %>%
  pack_rows("Hispanic", 55, 59) %>%
  pack_rows("Multiracial", 60, 64) %>%
  pack_rows("Native Hawaiian/Pacific Islander", 65, 67) %>%
  pack_rows("White", 68, 72) %>%
  pack_rows("AIAN-White", 73, 76) %>%
  pack_rows("Asian-White", 77, 81) %>%
  pack_rows("Black-White", 82, 86) %>%
  pack_rows("Hispanic-White", 87, 91) %>%
  pack_rows("Multi-White", 92, 96) %>%
  pack_rows("NHPI-White", 97, 99)

table_file6 = here("results", "aapc_all_results.rds")
saveRDS(all_aapc, file = table_file6)
```

## Plot of US Race Rate Ratios (White=ref)

```{r, echo=FALSE, message=FALSE}
US_jpoutput_raceratios_aapc<-jpoutput_raceratios_aapc %>% filter(geo %in% c("United States"))
fig_US_raceratio_AAPC<-ggplot(US_jpoutput_raceratios_aapc, aes(type, AAPC)) +   
      geom_point() + theme_minimal() + ggtitle("US") + theme(plot.title = element_text(hjust = 0.5)) +
      geom_errorbar(aes(ymin = AAPC_lCI, ymax = AAPC_uCI)) +
      scale_x_discrete(labels=c('AIAN', 'Asian', 'Black', 'Hispanic', 'Multi', 'NHPI')) +
      xlab("Race Ratio (vs White)") + ylab("Avg Annual % Change")

South_jpoutput_raceratios_aapc<-jpoutput_raceratios_aapc %>% filter(geo %in% c("South"))
fig_South_raceratio_AAPC<-ggplot(South_jpoutput_raceratios_aapc, aes(type, AAPC)) +   
      geom_point() + theme_minimal() + ggtitle("South") + theme(plot.title = element_text(hjust = 0.5)) +
      geom_errorbar(aes(ymin = AAPC_lCI, ymax = AAPC_uCI)) +
      scale_x_discrete(labels=c('AIAN', 'Asian', 'Black', 'Hispanic', 'Multi', 'NHPI')) +
      xlab("Race Ratio (vs White)") + ylab("Avg Annual % Change")

Northeast_jpoutput_raceratios_aapc<-jpoutput_raceratios_aapc %>% filter(geo %in% c("Northeast"))
fig_Northeast_raceratio_AAPC<-ggplot(Northeast_jpoutput_raceratios_aapc, aes(type, AAPC)) +        
      geom_point() + theme_minimal() + ggtitle("Northeast") + theme(plot.title = element_text(hjust = 0.5)) +
      geom_errorbar(aes(ymin = AAPC_lCI, ymax = AAPC_uCI)) +
      scale_x_discrete(labels=c('Asian', 'Black', 'Hispanic', 'Multi')) +
      xlab("Race Ratio (vs White)") + ylab("Avg Annual % Change")

West_jpoutput_raceratios_aapc<-jpoutput_raceratios_aapc %>% filter(geo %in% c("West"))
fig_West_raceratio_AAPC<-ggplot(West_jpoutput_raceratios_aapc, aes(type, AAPC)) + 
      geom_point() + theme_minimal() + ggtitle("West") + theme(plot.title = element_text(hjust = 0.5)) +
      geom_errorbar(aes(ymin = AAPC_lCI, ymax = AAPC_uCI)) +
      scale_x_discrete(labels=c('AIAN', 'Asian', 'Black', 'Hispanic', 'Multi', 'NHPI')) +
      xlab("Race Ratio (vs White)") + ylab("Avg Annual % Change")

Midwest_jpoutput_raceratios_aapc<-jpoutput_raceratios_aapc %>% filter(geo %in% c("Midwest"))
fig_Midwest_raceratio_AAPC<-ggplot(Midwest_jpoutput_raceratios_aapc, aes(type, AAPC)) + 
      geom_point() + theme_minimal() + ggtitle("Midwest") + theme(plot.title = element_text(hjust = 0.5)) +
      geom_errorbar(aes(ymin = AAPC_lCI, ymax = AAPC_uCI)) +
      scale_x_discrete(labels=c('AIAN', 'Asian', 'Black', 'Hispanic', 'Multi')) +
      xlab("Race Ratio (vs White)") + ylab("Avg Annual % Change")

rr_reg<-grid.arrange(arrangeGrob(fig_Northeast_raceratio_AAPC, fig_South_raceratio_AAPC, fig_Midwest_raceratio_AAPC, fig_West_raceratio_AAPC, nrow = 2))

rr_reg
figure_file = here("results","raceratio_byregion.png")
ggsave(filename = figure_file, plot=rr_reg) 
```

## Plot of US Rates by Race

```{r, echo=FALSE, message=FALSE}
US_jpoutput_race<-jpoutput_race_data %>% filter(geo %in% c("United States"))
US_race_rate_fig <- ggplot(US_jpoutput_race, aes(x=year, y=log(rate), size=log(rate), color=type)) + theme_light() +
    geom_point(alpha=0.5) +
    scale_y_continuous("Log Rate") +
    scale_x_continuous("Year", breaks=c(2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) + ggtitle("US Rates by Race") + 
    theme(plot.title = element_text(hjust = 0.5)) + guides(color=guide_legend(title="Race/Ethnicity"), size=guide_legend(title="Log Rate"))

US_race_rate_fig
figure_file = here("results","racerates_allUS.png")
ggsave(filename = figure_file, plot=US_race_rate_fig) 
```

## Plot of Regional Rates by Race

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Northeast
Northeast_jpoutput_race<-jpoutput_race_data %>% filter(geo %in% c("Northeast"))
Northeast_race_rate_fig <- ggplot(Northeast_jpoutput_race, aes(x=year, y=log(rate), size=log(rate), color=type)) + theme_light() +
    geom_point(alpha=0.5) +
    scale_y_continuous("Log Rate") +
    scale_x_continuous("Year", breaks=c(2008,2010,2012,2014,2016,2018,2020)) + ggtitle("Northeast") + 
    theme(plot.title = element_text(hjust = 0.5)) + guides(color=guide_legend(title="Race/Ethnicity"), size=guide_legend(title="Log Rate"))

# South
South_jpoutput_race<-jpoutput_race_data %>% filter(geo %in% c("South"))
South_race_rate_fig <- ggplot(South_jpoutput_race, aes(x=year, y=log(rate), size=log(rate), color=type)) + theme_light() +
    geom_point(alpha=0.5) +
    scale_y_continuous("Log Rate") +
    scale_x_continuous("Year", breaks=c(2008,2010,2012,2014,2016,2018,2020)) + ggtitle("South") + 
    theme(plot.title = element_text(hjust = 0.5)) + guides(color=guide_legend(title="Race/Ethnicity"), size=guide_legend(title="Log Rate"))

# Midwest
Midwest_jpoutput_race<-jpoutput_race_data %>% filter(geo %in% c("Midwest"))
Midwest_race_rate_fig <- ggplot(Midwest_jpoutput_race, aes(x=year, y=log(rate), size=log(rate), color=type)) + theme_light() +
    geom_point(alpha=0.5) +
    scale_y_continuous("Log Rate") +
    scale_x_continuous("Year", breaks=c(2008,2010,2012,2014,2016,2018,2020)) + ggtitle("Midwest") + 
    theme(plot.title = element_text(hjust = 0.5)) + guides(color=guide_legend(title="Race/Ethnicity"), size=guide_legend(title="Log Rate"))

# West
West_jpoutput_race<-jpoutput_race_data %>% filter(geo %in% c("West"))
West_race_rate_fig <- ggplot(West_jpoutput_race, aes(x=year, y=log(rate), size=log(rate), color=type)) + theme_light() +
    geom_point(alpha=0.5) +
    scale_y_continuous("Log Rate") +
    scale_x_continuous("Year", breaks=c(2008,2010,2012,2014,2016,2018,2020)) + ggtitle("West") + 
    theme(plot.title = element_text(hjust = 0.5)) + guides(color=guide_legend(title="Race/Ethnicity"), size=guide_legend(title="Log Rate"))

r_reg<-ggarrange(Northeast_race_rate_fig, South_race_rate_fig, Midwest_race_rate_fig, West_race_rate_fig, ncol=2, nrow = 2, common.legend = TRUE, legend="right")

r_reg
figure_file = here("results","racerates_region.png")
ggsave(filename = figure_file, plot=r_reg) 
```

## Joinpoint Plots for Race

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Black
US_jpoutput_black<-jpoutput_race_data %>% filter(geo %in% c("United States") & type %in% c("Black"))
black_jp_plot <- ggplot(US_jpoutput_black, aes(x=year, y=rate)) + geom_point() + 
  theme_light() + geom_line(aes(year, as.numeric(jp)), color="blue") +
  scale_y_continuous("Rate", breaks = c(0, 20, 40, 60, 80), limits=c(0, 80)) +
  scale_x_continuous("Year", breaks=c(2008,2011,2014,2017,2020)) + ggtitle("All US: Black") + theme(plot.title = element_text(hjust = 0.5))

# White
US_jpoutput_white<-jpoutput_race_data %>% filter(geo %in% c("United States") & type %in% c("White"))
white_jp_plot <- ggplot(US_jpoutput_white, aes(x=year, y=rate)) + geom_point() + 
  theme_light() + geom_line(aes(year, as.numeric(jp)), color="blue") +
  scale_y_continuous("Rate", breaks = c(0, 20, 40, 60, 80), limits=c(0, 80)) +
  scale_x_continuous("Year", breaks=c(2008,2011,2014,2017,2020)) + ggtitle("All US: White") + theme(plot.title = element_text(hjust = 0.5))

# Hispanic
US_jpoutput_hisp<-jpoutput_race_data %>% filter(geo %in% c("United States") & type %in% c("Hispanic"))
hisp_jp_plot <- ggplot(US_jpoutput_hisp, aes(x=year, y=rate)) + geom_point() + 
  theme_light() + geom_line(aes(year, as.numeric(jp)), color="blue") +
  scale_y_continuous("Rate", breaks = c(0, 20, 40, 60, 80), limits=c(0, 80)) +
  scale_x_continuous("Year", breaks=c(2008,2011,2014,2017,2020)) + ggtitle("All US: Hispanic") + theme(plot.title = element_text(hjust = 0.5))

jp_all_bwh<-grid.arrange(arrangeGrob(black_jp_plot, hisp_jp_plot, white_jp_plot, nrow = 1))

jp_all_bwh
figure_file = here("results","raceBWH_joinpoints_allUS.png")
ggsave(filename = figure_file, plot=jp_all_bwh) 
```

## Supplemental Table of APCs

```{r, echo=FALSE, message=FALSE}
jpoutput_overall_apc$Type<-"Overall" 
jpoutput_overall_apc <- jpoutput_overall_apc[, c(12, 1:11)]
all_apc<-rbind(jpoutput_overall_apc,jpoutput_sex_apc,jpoutput_age_apc,jpoutput_race_apc,jpoutput_raceratio_apc)

kable(all_apc,caption = "All APC Results", align="rc") %>% 
  row_spec(row = 0, color = "white", background = "#004987") %>% kable_styling(full_width=T, bootstrap_options = "striped", font_size=10) %>% 
  pack_rows("Overall", 1, 15) %>%
  pack_rows("Female", 16, 24) %>%
  pack_rows("Male", 25, 32) %>%
  pack_rows("Age 13-24", 33, 41) %>%
  pack_rows("Age 25-34", 42, 54) %>%
  pack_rows("Age 35-44", 55, 63)  %>%
  pack_rows("Age 45-54", 64, 70)  %>%
  pack_rows("Age 55+", 71, 76)  %>%
  pack_rows("American Indian/Alaska Native", 77, 80) %>%
  pack_rows("Asian", 81, 90) %>%
  pack_rows("Black", 91, 102) %>%
  pack_rows("Hispanic", 103, 113) %>%
  pack_rows("Multiracial", 114, 120) %>%
  pack_rows("Native Hawaiian/Pacific Islander", 121, 123) %>%
  pack_rows("White", 124, 130) %>%
  pack_rows("AIAN-White", 131, 134) %>%
  pack_rows("Asian-White", 135, 142) %>%
  pack_rows("Black-White", 143, 147) %>%
  pack_rows("Hispanic-White", 148, 156) %>%
  pack_rows("Multi-White", 157, 165) %>%
  pack_rows("NHPI-White", 166, 168)

table_file7 = here("results", "apc_all_results.rds")
saveRDS(all_apc, file = table_file7)
```

## Rate difference comparisons by race (Black and Hispanic vs White)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ratediff_us_compare_bwh<-alldata %>% filter(Geo %in% c("United States")) %>%  dplyr::select(Year,Black_Rate,White_Rate,Hispanic_Rate, Asian_Rate, AIAN_Rate, MultRace_Rate,NHPI_Rate)
ratediff_us_compare_bwh$DiffWB<-ratediff_us_compare_bwh$White_Rate - ratediff_us_compare_bwh$Black_Rate
ratediff_us_compare_bwh$DiffWH<-ratediff_us_compare_bwh$White_Rate - ratediff_us_compare_bwh$Hispanic_Rate
nat_diff_bwh <- plot_ly(ratediff_us_compare_bwh, x = ~Year, y = ~DiffWB, name = "White-Black Rate Difference", type="scatter", mode = "lines", line = list(width = 3, dash = "solid", color = "blue")) 
nat_diff_bwh <- nat_diff_bwh %>% add_trace(y = ~DiffWH, name = "White-Hispanic Rate Difference", mode = 'lines', line = list(width = 3, dash = "solid", color = "green"))
nat_diff_bwh <- nat_diff_bwh %>% layout(title="New Diagnoses Rate Differences by Year, all US", legend=list(title=list(text=' ')), 
                  xaxis = list(title = 'Year'),
                  yaxis = list(title = 'Rate Difference'))

nat_diff_bwh
plotly::export(p=nat_diff_bwh, file = here("results", "nat_diff_bwh.png"))
```

## Supplemental Table of Rates, National

```{r, echo=FALSE, message=FALSE}
national<-alldata %>% filter(Geo %in% c("United States")) %>% dplyr::select(Year,Overall_Rate,Male_Rate,Female_Rate,Black_Rate,White_Rate,Hispanic_Rate,Asian_Rate,AIAN_Rate,MultRace_Rate,NHPI_Rate,Age13t24_Rate,Age25t34_Rate,Age35t44_Rate,Age45t54_Rate,Age55p_Rate)
national %>% gt::gt() %>% 
  gt::tab_header(title = "National Rate Data") %>%
  gt::tab_style(style = list(gt::cell_fill(color = "#7cb9e8"),
                             gt::cell_text(weight = "bold")),
                locations = gt::cells_body(columns = Year))

table_file8 = here("results", "all_us_rates.rds")
saveRDS(national, file = table_file8)
```

# Analyze State Race Data with Google Trend HIV Search

```{r}
state_race_gtrend<-readRDS(here("data", "processed_data", "state_race_gtrend.rds"))
gtrends_2021_2022 <- readxl::read_excel(here("data","raw_data", "Misc", "Gtrends_2021_2022.xlsx")) 
p_2022 <- gtrends_2021_2022 %>% filter(Year %in% c(2022))

linear_reg() %>% 
  set_engine("keras")
lm_mod <- linear_reg()

# predict 2022 black-white
lm_fit <- 
  lm_mod %>% 
  fit(black_white_rateratio ~ Gtrend + Year + State , data = state_race_gtrend)
lm_fit
tidy(lm_fit)
tidy(lm_fit) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))

mean_pred <- predict(lm_fit, new_data = p_2022)
conf_int_pred <- predict(lm_fit, new_data = p_2022, type = "conf_int")
glance(lm_fit) %>%
  dplyr::select(adj.r.squared, sigma, AIC, BIC, p.value)

plot_data <- 
  p_2022 %>% 
  bind_cols(mean_pred) %>% 
  bind_cols(conf_int_pred)

compare <- state_race_gtrend  %>% filter(Year %in% c(2020)) %>% select(State, black_white_rateratio, hispanic_white_rateratio) 
plot_data_c <- merge(plot_data,compare,by=c("State"),all=T)

south<-plot_data_c %>% filter(State %in% c('Alabama','Arkansas','Delaware','District of Columbia','Florida','Georgia',
                                         'Kentucky','Louisiana','Maryland','Mississippi','North Carolina','Oklahoma',
                                         'South Carolina','Tennessee','Texas','Virginia','West Virginia'))
northeast<-plot_data_c %>% filter(State %in% c('Connecticut', 'Maine', 'Massachusetts', 'New Hampshire', 'New Jersey', 'New York', 
                                             'Pennsylvania', 'Rhode Island', 'Vermont'))
west<-plot_data_c %>% filter(State %in% c('Alaska', 'Arizona', 'California', 'Colorado', 'Hawaii', 'Idaho', 'Montana', 'Nevada', 'New Mexico',
                                        'Oregon', 'Utah', 'Washington', 'Wyoming'))
midwest<-plot_data_c %>% filter(State %in% c('Illinois', 'Indiana', 'Iowa', 'Kansas', 'Michigan', 'Minnesota', 'Missouri', 'Nebraska', 'North Dakota',
                                           'Ohio','South Dakota', 'Wisconsin'))

south22<-ggplot(south, aes(x = State)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "Black-White Rate-Ratio") + 
  scale_x_discrete(labels=c('AL','AR','DE','DC','FL','GA', 'KY','LA','MD','MS','NC','OK',
                                         'SC','TN','TX','VA','WV')) +
  ggtitle(label="SOUTH: Predict 2022") + 
  theme_light() + 
  geom_point(aes(y = black_white_rateratio, color="red")) + 
  theme(legend.title = element_blank()) + 
  scale_colour_discrete(labels=c('Actual Black-White Rate Ratio, 2020'))

northeast22<-ggplot(northeast, aes(x = State)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "Black-White Rate-Ratio") + 
  scale_x_discrete(labels=c('CT', 'ME', 'MA', 'NH', 'NJ', 'NY', 'PA', 'RI', 'VT')) +
  ggtitle(label="NORTHEAST: Predict 2022") + 
  theme_light() + 
  geom_point(aes(y = black_white_rateratio, color="red")) + 
  theme(legend.title = element_blank()) + 
  scale_colour_discrete(labels=c('Actual Black-White Rate Ratio, 2020'))

west22<-ggplot(west, aes(x = State)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "Black-White Rate-Ratio") + 
  scale_x_discrete(labels=c('AK', 'AZ', 'CA', 'CO', 'HI', 'ID', 'MT', 'NV', 'NM', 
                            'OR', 'UT', 'WA', 'WY')) +
  ggtitle(label="WEST: Predict 2022") + 
  theme_light() + 
  geom_point(aes(y = black_white_rateratio, color="red")) + 
  theme(legend.title = element_blank()) + 
  scale_colour_discrete(labels=c('Actual Black-White Rate Ratio, 2020'))

midwest22<-ggplot(midwest, aes(x = State)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "Black-White Rate-Ratio") + 
  scale_x_discrete(labels=c('IL', 'IN', 'IA', 'KS', 'MI', 'MN', 'MO', 'NE', 'ND',
                            'OH','SD', 'WI')) +
  ggtitle(label="MIDWEST: Predict 2022") + 
  theme_light() + 
  geom_point(aes(y = black_white_rateratio, color="red")) + 
  theme(legend.title = element_blank()) + 
  scale_colour_discrete(labels=c('Actual Black-White Rate Ratio, 2020'))

predict_2022_blackwhite<-ggarrange(northeast22, south22, west22, midwest22, ncol=2, nrow = 2, common.legend = TRUE, legend="bottom")
predict_2022_blackwhite
figure_file = here("results","Predict2022_BWratio.png")
ggsave(filename = figure_file, plot=predict_2022_blackwhite)

# predict 2022 hispanic-white
lm_fit <- 
  lm_mod %>% 
  fit(hispanic_white_rateratio ~ Gtrend + Year + State , data = state_race_gtrend)
lm_fit
tidy(lm_fit)
tidy(lm_fit) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2)) 
mean_pred <- predict(lm_fit, new_data = p_2022)
conf_int_pred <- predict(lm_fit, new_data = p_2022, type = "conf_int")
glance(lm_fit) %>%
  dplyr::select(adj.r.squared, sigma, AIC, BIC, p.value)

plot_data <- 
  p_2022 %>% 
  bind_cols(mean_pred) %>% 
  bind_cols(conf_int_pred)

compare <- state_race_gtrend  %>% filter(Year %in% c(2020)) %>% select(State, black_white_rateratio, hispanic_white_rateratio) 
plot_data_c <- merge(plot_data,compare,by=c("State"),all=T)

south<-plot_data_c %>% filter(State %in% c('Alabama','Arkansas','Delaware','District of Columbia','Florida','Georgia',
                                         'Kentucky','Louisiana','Maryland','Mississippi','North Carolina','Oklahoma',
                                         'South Carolina','Tennessee','Texas','Virginia','West Virginia'))
northeast<-plot_data_c %>% filter(State %in% c('Connecticut', 'Maine', 'Massachusetts', 'New Hampshire', 'New Jersey', 'New York', 
                                             'Pennsylvania', 'Rhode Island', 'Vermont'))
west<-plot_data_c %>% filter(State %in% c('Alaska', 'Arizona', 'California', 'Colorado', 'Hawaii', 'Idaho', 'Montana', 'Nevada', 'New Mexico',
                                        'Oregon', 'Utah', 'Washington', 'Wyoming'))
midwest<-plot_data_c %>% filter(State %in% c('Illinois', 'Indiana', 'Iowa', 'Kansas', 'Michigan', 'Minnesota', 'Missouri', 'Nebraska', 'North Dakota',
                                           'Ohio','South Dakota', 'Wisconsin'))

south22<-ggplot(south, aes(x = State)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "Hispanic-White Rate-Ratio") + 
  scale_x_discrete(labels=c('AL','AR','DE','DC','FL','GA', 'KY','LA','MD','MS','NC','OK',
                                         'SC','TN','TX','VA','WV')) +
  ggtitle(label="SOUTH: Predict 2022") + 
  theme_light() + 
  geom_point(aes(y = hispanic_white_rateratio, color="red")) + 
  theme(legend.title = element_blank()) + 
  scale_colour_discrete(labels=c('Actual Hispanic-White Rate Ratio, 2020'))

northeast22<-ggplot(northeast, aes(x = State)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "Hispanic-White Rate-Ratio") + 
  scale_x_discrete(labels=c('CT', 'ME', 'MA', 'NH', 'NJ', 'NY', 'PA', 'RI', 'VT')) +
  ggtitle(label="NORTHEAST: Predict 2022") + 
  theme_light() + 
  geom_point(aes(y = hispanic_white_rateratio, color="red")) + 
  theme(legend.title = element_blank()) + 
  scale_colour_discrete(labels=c('Actual Hispanic-White Rate Ratio, 2020'))

west22<-ggplot(west, aes(x = State)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "Hispanic-White Rate-Ratio") + 
  scale_x_discrete(labels=c('AK', 'AZ', 'CA', 'CO', 'HI', 'ID', 'MT', 'NV', 'NM', 
                            'OR', 'UT', 'WA', 'WY')) +
  ggtitle(label="WEST: Predict 2022") + 
  theme_light() + 
  geom_point(aes(y = hispanic_white_rateratio, color="red")) + 
  theme(legend.title = element_blank()) + 
  scale_colour_discrete(labels=c('Actual Hispanic-White Rate Ratio, 2020'))

midwest22<-ggplot(midwest, aes(x = State)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "Hispanic-White Rate-Ratio") + 
  scale_x_discrete(labels=c('IL', 'IN', 'IA', 'KS', 'MI', 'MN', 'MO', 'NE', 'ND',
                            'OH','SD', 'WI')) +
  ggtitle(label="MIDWEST: Predict 2022") + 
  theme_light() + 
  geom_point(aes(y = hispanic_white_rateratio, color="red")) + 
  theme(legend.title = element_blank()) + 
  scale_colour_discrete(labels=c('Actual Hispanic-White Rate Ratio, 2020'))

predict_2022_hispanicwhite<-ggarrange(northeast22, south22, west22, midwest22, ncol=2, nrow = 2, common.legend = TRUE, legend="bottom")
predict_2022_hispanicwhite
figure_file = here("results","Predict2022_HWratio.png")
ggsave(filename = figure_file, plot=predict_2022_hispanicwhite) 
```